
\begin{glossary}

\rdv{We can just copy a big chunk of this from the FL MOOC.}

\term{Absolute Zero}{
The lowest temperature possible, equivalent to -273.15$^{\deg}$C (or
0$^{\deg}$ on the
absolute Kelvin scale), at which point atoms cease to move altogether
and molecular energy is minimal. The idea that it is impossible,
through any physical process, to lower the temperature of a system to
zero is known as the Third Law of Thermodynamics.}


\term{Alpha Particle (Alpha Decay)}{
A particle of 2 protons and 2 neutrons (essentially a helium nucleus)
that is emitted by an unstable radioactive nucleus during radioactive
decay. It is a relatively low-penetration particle due its
comparatively low energy and high mass.}

\term{Angular Momentum}{A measure of the momentum of a body in rotational
motion about its centre of mass. Technically, the angular momentum of
a body is equal to the mass of the body multiplied by the cross
product of the position vector of the particle with its velocity
vector. The angular momentum of a system is the sum of the angular
momenta of its constituent particles, and this total is conserved
unless acted on by an outside force.}

\term{Anthropic Principle}
{The idea that the fundamental constants of physics and chemistry are
just right (or ``fine-tuned'') to allow the universe and life as we know
it to exist, and indeed that the universe is only as it is because we
are here to observe it. Thus, we find ourselves in the kind of
universe, and on the kind of planet, where conditions are ripe for our
form of life.}

\term{Antimatter}{Pair production and pair annihilation
of hydrogen and antihydrogen particles. 
A large
accumulation of antiparticles---antiprotons, antineutrons and
positrons (antielectrons)---which have opposite properties to normal
particles (e.g. electrical charge), and which can come together to
make antiatoms. When matter and antimatter meet, they self-destruct in
a burst of high-energy photons or gamma rays. The laws of physics seem
to predict a pretty much 50/50 mix of matter and antimatter, despite
the observable universe apparently consisting almost entirely of
matter, known as the ``baryon asymmetry problem''.}

%(Source: Riken Research:
%\url{http://www.rikenresearch.riken.jp/eng/frontline/5444})\\ 

\end{glossary}