
\begin{partintro}
\partintrotitle{Introduction to the third lesson block}
With the basics in place, in this block of lessons we introduce quantum teleportation and two forms of QKD (single-photon BB84 and entanglement-based E91) as three of the intermediate-layer building blocks.  It is worth noting that teleportation and QKD are often described as "applications" of the Quantum Internet. In our terminology, however, teleportation is not an application protocol; it is instead a function that is used within systems to build more complete systems. In classical Internet terms, it is very roughly like the transport protocol TCP, whose job it is to see that data is correctly transmitted end-to-end but does not care about how that data is used. Even this analogy, however, is poor; teleportation does not assess or correct errors. For a more complete, robust system, then, it must be combined with an error correction or at least error detection method, as we will see in Sec.~\ref{sec:purification} in the following block of lessons when we study purification.

QKD likewise provides a function that is of little use alone, but must be incorporated into larger systems to provide a service that is used by end users. QKD, whether single-photon or entanglement-based, requires the measuremet of qubits and the collapse of superposition and/or entanglement. This bridging of the quantum and classical levels of a complete system is of particular interest, but the details of how the random numbers generated in QKD are incorporated into a complete system are deferred to Sec.~\ref{sec:classical-integration}.
\end{partintro}
